Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-14 paper/manuscript.Rmd
re-align paragraphs
minor edits

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and to access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, which have hollow hair
shafts that aid in flotation, the energetic expenditure of swimming for
quadrupedal mammals is significantly higher than walking or running, and
drowning is also possible [@Miller1985]. Despite the abundance of coastal and
island caribou herds, only a handful of studies have documented caribou swimming
in the ocean, and most of these lack observation, but rather deduce swimming
based on presence of caribou on a previously uninhabitated island  [e.g.
@Miller1995; @Miller2002; @Jeffery2007; @Ricca2012]. Caribou may swim between
islands in the ocean for similar reasons they swim in freshwater [@Leblond2016],
mainly concerning predator avoidance or to access new habitats and foraging opportunities. We propose
that one such explanation, the forage limitation hypothesis, is the most likely
explanation for oceanic swimming for caribou that live on islands in the Fogo
Island archipelago.

@@

1) How do the energetic cost and risk of swimming differ from walking on sea
ice? Walking over solid ice is likely less energetically expensive than swimming, but may pose a greater risk of falling through depending on the thickness and stability of ice. What proportion of inter-island movements are swimming events compared to walking on ice? The presence and extent of sea ice varies annually in our study area and we are unable to determine the exact timing or duration of sea ice from past years. However, we qualitatively delineate the timing of typical sea ice arrival and departure (Figure 2B) and deduce that although the majority of inter-island movements occur between April 1 and December 31, some inter-island movements appear to be caribou walking on sea ice between islands. Future research on the trade-offs of these types of movement should consider both how the costs associated with each differ, but also how the potential benefits - foraging opportunities - may vary seasonally.

>>>>

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and to access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, which have hollow hair
shafts that aid in flotation, the energetic expenditure of swimming for
quadrupedal mammals is significantly higher than walking or running, and
drowning is also possible [@Miller1985]. Despite the abundance of coastal and
island caribou herds, only a few studies have documented caribou swimming
in the ocean, and most of these lack observation, but rather deduce swimming
based on presence of caribou on a previously uninhabitated island  [e.g.
@Miller1995; @Miller2002; @Jeffery2007; @Ricca2012]. Caribou may swim between
islands in the ocean for similar reasons they swim in freshwater [@Leblond2016],
mainly concerning predator avoidance or to access new habitats and foraging
opportunities. We propose that one such explanation, the forage limitation
hypothesis, is the most likely explanation for oceanic swimming for caribou that
live on islands in the Fogo Island archipelago.

@@

1) How do the energetic cost and risk of swimming differ from walking on sea
ice? Walking over solid ice is likely less energetically expensive than
swimming, but may pose a greater risk of falling through depending on the
thickness and stability of ice. What proportion of inter-island movements are
swimming events compared to walking on ice? Although the presence and extent of sea ice
varies annually,  we
deduce that although the majority of inter-island movements occur between April
1 and December 31, some inter-island movements appear to be caribou walking on
sea ice between islands (Figure 2B). Future research on the trade-offs of these types of
movement should consider both how the costs associated with each differ, but
also how the potential benefits - foraging opportunities - may vary seasonally.
